A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol

### 2023-04-14 paper.Rmd
section 2.2

Figure \ref{fig:residual-plot-common-departures}D) shows a scatterplot where the residuals have a skewed distribution, as seen by the uneven vertical spread. Unlike non-linearity and heteroskedasticity, non-normality is usually detected with a different type of residual plot, a histogram or normal probability plot.  Because we focus on scatterplots, non-normality is not one of the departures examined in this paper.

@@

We would expect the RESET test for non-linearity to reject residual plot B, the BP test for heteroskedasticity to reject the residual plot C, and SW test for non-normality to reject residual plot D, which they all do and correctly fail to reject residual plot A. Interestingly, the BP and SW tests also reject the residual plots exhibiting structure that they weren't designed for. @cook1982residuals discusses that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures, that is, correctly rejected but for the wrong reason, a phenomenon known as the "Type III error". Also, some types of departure can also have elements of other types of departure, for example, non-linearity could be viewed as heteroskedasticity. Additionally, other problesms such as outliers can trigger the rejection [@cook_applied_1999].

With large sample sizes, hypothesis tests are may reject a residual plot when there is only a slight departure (XXX is there a reference for this statement?). While such rejections may be statistically correct, their sensitivity may render the results impractical. Therefore, a key goal of residual plot diagnostics is to identify potential issues that could lead to incorrect conclusions or errors in subsequent analyses. However, minor defects in the model are unlikely to have a significant impact and may be best disregarded for practical purposes. In fact, the experiment discussed in this paper specifically addresses this.

>>>>

Figure \ref{fig:residual-plot-common-departures}D) shows a scatterplot where the residuals have a skewed distribution, as seen by the uneven vertical spread. Unlike non-linearity and heteroskedasticity, non-normality is usually detected with a different type of residual plot, a histogram or normal probability plot.  Because we focus on scatterplots, non-normality is not one of the departures examined in depth in this paper.

@@

We would expect the RESET test for non-linearity to reject residual plot B, the BP test for heteroskedasticity to reject the residual plot C, and SW test for non-normality to reject residual plot D, which they all do and correctly fail to reject residual plot A. Interestingly, the BP and SW tests also reject the residual plots exhibiting structure that they weren't designed for. @cook1982residuals discusses that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures. This could be considered a Type III error, where the null hypothesis of good residuals is correctly rejected but for the wrong reason. Also, some types of departure can have elements of other types of departure, for example, non-linearity could be viewed as heteroskedasticity. Additionally, other problesms such as outliers can trigger the rejection [@cook_applied_1999].

With large sample sizes, hypothesis tests may reject the null hypothesis when there is only a small effect. (This is common knowledge but a good discussion can be found in @kirk1996.) While such rejections may be statistically correct, their sensitivity may render the results impractical. This is especially pertinent for residual diagnostics. Therefore, a key goal of residual plot diagnostics is to identify potential issues that could lead to incorrect conclusions or errors in subsequent analyses. However, minor defects in the model are unlikely to have a significant impact and may be best disregarded for practical purposes. In fact, the experiment discussed in this paper specifically addresses this.

### 2023-04-14 paper.tex
section 2.2

We would expect the RESET test for non-linearity to reject residual plot
B, the BP test for heteroskedasticity to reject the residual plot C, and
SW test for non-normality to reject residual plot D, which they all do
and correctly fail to reject residual plot A. Interestingly, the BP and
SW tests also reject the residual plots exhibiting structure that they
weren't designed for. \citet{cook1982residuals} discusses that most
residual-based tests for particular types of departure from model
assumptions are also sensitive to other types of departures, that is,
correctly rejected but for the wrong reason, a phenomenon known as the
``Type III error''. Also, some types of departure can also have elements
of other types of departure, for example, non-linearity could be viewed
as heteroskedasticity. Additionally, other problesms such as outliers
can trigger the rejection \citep{cook_applied_1999}.

With large sample sizes, hypothesis tests are may reject a residual plot
when there is only a slight departure (XXX is there a reference for this
statement?). While such rejections may be statistically correct, their
sensitivity may render the results impractical. Therefore, a key goal of
residual plot diagnostics is to identify potential issues that could
lead to incorrect conclusions or errors in subsequent analyses. However,
minor defects in the model are unlikely to have a significant impact and
may be best disregarded for practical purposes. In fact, the experiment
discussed in this paper specifically addresses this.

>>>>

We would expect the RESET test for non-linearity to reject residual plot
B, the BP test for heteroskedasticity to reject the residual plot C, and
SW test for non-normality to reject residual plot D, which they all do
and correctly fail to reject residual plot A. Interestingly, the BP and
SW tests also reject the residual plots exhibiting structure that they
weren't designed for. \citet{cook1982residuals} discusses that most
residual-based tests for particular types of departure from model
assumptions are also sensitive to other types of departures. This could
be considered a Type III error, where the null hypothesis of good
residuals is correctly rejected but for the wrong reason. Also, some
types of departure can have elements of other types of departure, for
example, non-linearity could be viewed as heteroskedasticity.
Additionally, other problesms such as outliers can trigger the rejection
\citep{cook_applied_1999}.

With large sample sizes, hypothesis tests may reject the null hypothesis
when there is only a small effect. (This is common knowledge but a good
discussion can be found in \citet{kirk1996}.) While such rejections may
be statistically correct, their sensitivity may render the results
impractical. This is especially pertinent for residual diagnostics.
Therefore, a key goal of residual plot diagnostics is to identify
potential issues that could lead to incorrect conclusions or errors in
subsequent analyses. However, minor defects in the model are unlikely to
have a significant impact and may be best disregarded for practical
purposes. In fact, the experiment discussed in this paper specifically
addresses this.
